---
title: "Data Export"
subtitle: A dataset file exported from the pilot database
---

```{css, echo=FALSE}
th {
  white-space: nowrap;
}
```

The data file was created for the initial analysis using the MUDAB data.

## Data Availability

A tab-delimited file containing all major data from the database is available for download at the following location.

 * [pilot_mudab.tsv.gz](https://github.com/seafood-hazards/mudab-pilot/releases/download/v0.1.4/pilot_mudab.tsv.gz)

## Data File Summary

The file `pilot_mudab.tsv.gz` is a denormalized (flat) export of the relational database, designed to facilitate immediate analysis without requiring complex SQL joins.

## Column Definition

```{r setup, include=FALSE}
library(tibble)
```

```{r}
tribble(
  ~`Column Name`,                    ~`Source Table`,      ~`Type`,    ~`Description`,

  # --- Keys ---
  "station_no",                      "station",            "INTEGER",  "Surrogate key linking to the station table.",
  "measurement_time_id",             "measurement_time",   "INTEGER",  "Surrogate key linking to the measurement_time table.",
  "analysis_method_id",              "analysis_method",    "INTEGER",  "Surrogate key linking to the analysis_method table.",
  "reference_material_id",           "reference_material", "INTEGER",  "Surrogate key linking to the reference_material table.",
  "survey_id",                       "survey",             "INTEGER",  "Surrogate key linking to the survey table.",
  "sample_no",                       "sample",             "INTEGER",  "Surrogate key linking to the sample table.",
  "lod_id",                          "lod",                "INTEGER",  "Surrogate key linking to the lod table.",

  # --- Station metadata ---
  "organisation",                    "station",            "TEXT",     "Organisation type responsible for the station (e.g. NATIONAL, OSPAR, HELCOM).",
  "project_affiliation",             "station",            "TEXT",     "Monitoring programme affiliation (e.g. BLMP+).",
  "responsible_institute",           "station",            "TEXT",     "Institute responsible for the station.",
  "region",                          "station",            "TEXT",     "Sea region (e.g. North Sea, Baltic Sea).",
  "water_body_category",             "station",            "TEXT",     "Water body category code.",

  # --- Measurement time ---
  "year",                            "measurement_time",   "REAL",     "Year of measurement.",
  "survey_start",                    "survey",             "REAL",     "Survey start date (integer format YYYYMMDD).",
  "survey_end",                      "survey",             "REAL",     "Survey end date (integer format YYYYMMDD).",

  # --- Location ---
  "station_latitude",                "survey",             "REAL",     "Latitude of the monitoring station (decimal degrees).",
  "station_longitude",               "survey",             "REAL",     "Longitude of the monitoring station (decimal degrees).",
  "dist_to_coast",                   "survey",             "REAL",     "Distance from station to nearest coastline (km).",
  "est_country",                     "survey",             "TEXT",     "Estimated country derived from station coordinates.",
  "country_code",                    "survey",             "TEXT",     "ISO country code (e.g. DE).",
  "municipality",                    "survey",             "TEXT",     "Municipality name.",
  "sea_name",                        "survey",             "TEXT",     "Name of the sea or ocean.",

  # --- Sample metadata ---
  "sampling_method",                 "sample",             "TEXT",     "Method used to collect the sample (e.g. BC = box corer, VV = Van Veen grab).",
  "matrix",                          "sample",             "TEXT",     "Sediment matrix fraction code (e.g. SED20, SED63, SED2000).",
  "sediment_composition",            "sample",             "TEXT",     "Field description of the sediment composition (translated to English).",
  "sediment_no",                     "sediment",           "INTEGER",  "Sequential index for multiple sediment layers within a sample.",
  "layer_upper_boundary",            "sample",             "REAL",     "Upper depth boundary of the sampled sediment layer (cm).",
  "layer_lower_boundary",            "sample",             "REAL",     "Lower depth boundary of the sampled sediment layer (cm).",

  # --- Parameter ---
  "parameter",                       "parameter",          "TEXT",     "ICES parameter code (e.g. HG, CD, ZN).",
  "parameter_name",                  "parameter",          "TEXT",     "Full name of the parameter (e.g. mercury, cadmium).",
  "parameter_group",                 "parameter",          "TEXT",     "Parameter group code (e.g. I-MET for metals and metalloids).",
  "cas_number",                      "parameter",          "TEXT",     "Chemical Abstracts Service registry number.",

  # --- Measurement ---
  "measured_value",                  "sediment",           "REAL",     "Measured concentration or numerical result.",
  "unit",                            "sediment",           "TEXT",     "Unit of measurement (e.g. mg/kg, g/kg).",
  "data_qualifier",                  "sediment",           "TEXT",     "Data quality qualifier code.",

  # --- Analysis method ---
  "analytical_laboratory",           "analysis_method",    "TEXT",     "Code of the laboratory that performed the analysis.",
  "internal_qa_count",               "sediment",           "REAL",     "Number of measurements used for internal quality assurance.",
  "internal_qa_detection_limit",     "lod",                "REAL",     "Limit of detection value for this parameter measurement.",
  "recovery_rate",                   "sediment",           "REAL",     "Analytical recovery rate (%).",
  "internal_qa_quantification_limit","lod",                "REAL",     "Limit of quantification value for this parameter measurement.",
  "expanded_uncertainty_pct",        "lod",                "REAL",     "Expanded measurement uncertainty expressed as a percentage.",
  "uncertainty_method",              "lod",                "TEXT",     "Method used to determine the measurement uncertainty.",
  "measurement_method_code",         "analysis_method",    "TEXT",     "Analytical measurement method code (e.g. AAS-CV, ICP-OES).",
  "chemical_treatment",              "analysis_method",    "TEXT",     "Chemical treatment applied to the sample (e.g. HF-C, HNO).",
  "physical_treatment",              "analysis_method",    "TEXT",     "Physical pre-treatment applied to the sample (e.g. DFRZ = freeze-drying).",
  "measurement_basis",               "analysis_method",    "TEXT",     "Basis of measurement (e.g. D = dry weight).",
  "accreditation",                   "analysis_method",    "TEXT",     "Accreditation status of the analytical laboratory."
)

```

