---
title: "EFSA Submission v1"
subtitle: "Data export of trace elements for EFSA submission"
---

```{css, echo=FALSE}
th {
  white-space: nowrap;
}
```

The page presents the mapping between the pilot database fields (see the [DB Schema](db-schema.qmd) page) and the [EFSA Format v1](efsa-format.qmd) columns, and the resulting data export of trace elements in background sediments from the pilot database.

```{r}
library(dplyr)
library(tibble)

efsa_submission <- tribble(
  ~Column,            ~`Lookup`, ~Optionality, ~Table,             ~`Column Names`,                                    ~`Fixed Text`,      ~Missing,
  "sampleId",         NA,        "Mandatory",  "sample",           "sample_id",                                        NA,                 NA,
  "recId",            "Yes",     "Mandatory",  NA,                 NA,                                                 NA,                 NA,
  "envComp",          NA,        "Mandatory",  NA,                 NA,                                                 "marine sediment",  NA,
  "pristineLoc",      NA,        "Mandatory",  NA,                 NA,                                                 "Yes",              NA,
  "sampleLocGC",      NA,        "Mandatory",  "survey",           "station_latitude + station_longitude",             NA,                 NA,
  "sampleLocCM",      NA,        "Mandatory",  "survey",           "est_country + municipality",                       NA,                 NA,
  "sampleDate",       NA,        "Mandatory",  "measurement_time", "measurement_date",                                 NA,                 NA,
  "traceElText",      NA,        "Mandatory",  "parameter",        "parameter_name",                                  NA,                 NA,
  "traceElCode",      "Yes",     "Mandatory",  NA,                 NA,                                                 NA,                 NA,
  "specText",         NA,        "Mandatory",  "parameter",        'parameter_name + " (Total)"',                     NA,                 NA,
  "specCode",         "Yes",     "Mandatory",  NA,                 NA,                                                 NA,                 NA,
  "conc",             NA,        "Mandatory",  "sediment",         "measured_value",                                  NA,                 NA,
  "unitText",         NA,        "Mandatory",  "sediment",         "unit",                                            NA,                 NA,
  "unitCode",         "Yes",     "Mandatory",  NA,                 NA,                                                 NA,                 NA,
  "weightText",       NA,        "Mandatory",  "analysis_method",  "measurement_basis",                               NA,                 NA,
  "weightCode",       "Yes",     "Mandatory",  NA,                 NA,                                                 NA,                 NA,
  "methAnText",       NA,        "Mandatory",  "analysis_method",  "measurement_method_code",                         NA,                 NA,
  "methAnCode",       "Yes",     "Mandatory",  NA,                 NA,                                                 NA,                 NA,
  "LOD",              NA,        "Optional",   "lod",              "internal_qa_detection_limit",                     NA,                 NA,
  "LOQ",              NA,        "Optional",   "lod",              "internal_qa_quantification_limit",                NA,                 NA,
  "accLab",           NA,        "Optional",   "analysis_method",  "accreditation",                                   NA,                 NA,
  "phSed",            NA,        "Optional",   NA,                 NA,                                                 NA,                 "Yes",
  "TextureSedClay",   NA,        "Optional",   NA,                 NA,                                                 NA,                 "Yes",
  "TextureSedSilt",   NA,        "Optional",   NA,                 NA,                                                 NA,                 "Yes",
  "TextureSedSand",   NA,        "Optional",   NA,                 NA,                                                 NA,                 "Yes",
  "unitText2",        NA,        "Optional",   NA,                 NA,                                                 NA,                 "Yes",
  "unitCode2",        "Yes",     "Optional",   NA,                 NA,                                                 NA,                 NA,
  "weightText2",      "Yes",     "Optional",   NA,                 NA,                                                 NA,                 NA,
  "weightCode2",      "Yes",     "Optional",   NA,                 NA,                                                 NA,                 NA,
  "ocSed",            NA,        "Optional",   NA,                 NA,                                                 NA,                 "Yes",
  "unitText3",        NA,        "Optional",   NA,                 NA,                                                 NA,                 "Yes",
  "unitCode3",        "Yes",     "Optional",   NA,                 NA,                                                 NA,                 NA,
  "weightText3",      "Yes",     "Optional",   NA,                 NA,                                                 NA,                 NA,
  "weightCode3",      "Yes",     "Optional",   NA,                 NA,                                                 NA,                 NA,
  "phWater",          NA,        "Optional",   NA,                 NA,                                                 NA,                 "Yes",
  "hardWater",        NA,        "Optional",   NA,                 NA,                                                 NA,                 "Yes",
  "DOC",              NA,        "Optional",   NA,                 NA,                                                 NA,                 "Yes",
  "unitText5",        NA,        "Optional",   NA,                 NA,                                                 NA,                 NA,
  "unitCode5",        "Yes",     "Optional",   NA,                 NA,                                                 NA,                 NA,
  "publicData",       NA,        "Optional",   NA,                 NA,                                                 "Yes",              NA,
  "refPublication",   NA,        "Optional",   NA,                 NA,                                                 NA,                 "Yes",
  "comments",         NA,        "Optional",   NA,                 NA,                                                 NA,                 "Yes",
  "Missing recId",        "Yes", "Optional",   NA,                 NA,                        NA,                 NA,
  "Missing envComp",      "Yes", "Optional",   NA,                 NA,                        NA,                 NA,
  "Missing pristineLoc",  "Yes", "Optional",   NA,                 NA,                        NA,                 NA,
  "Missing sampleLoc",    "Yes", "Optional",   NA,                 NA,                        NA,                 NA,
  "Missing sampleDate",   "Yes", "Optional",   NA,                 NA,                        NA,                 NA,
  "Missing traceEl",      "Yes", "Optional",   NA,                 NA,                        NA,                 NA,
  "Missing spec",         "Yes", "Optional",   NA,                 NA,                        NA,                 NA,
  "Missing conc",         "Yes", "Optional",   NA,                 NA,                        NA,                 NA,
  "Missing methAn",       "Yes", "Optional",   NA,                 NA,                        NA,                 NA
)

efsa_submission_full <- efsa_submission %>%
  mutate(
    across(
      where(is.character),
      ~coalesce(.x, "")
    )
  ) %>%
  rename(c(`Source Table` = "Table",
           `Source Columns` = "Column Names"))

efsa_submission <- efsa_submission_full %>%
  filter(Lookup != "Yes") %>%
  dplyr::select(-c(Lookup, Missing))

```

## Data Mapping for Submission
```{r}
efsa_submission
```

## Full Data Mapping
```{r}
efsa_submission_full
```

## Data Export Method
As the EFSA submission tool is an Excel file containing embedded macros, the data must be mapped column by column rather than through a standard row‑based export. This approach ensures that the structure of the submission file is not disrupted.
