---
title: "EFSA Submission v2"
subtitle: "Data export mapping for the FHF marine sediment data model"
---

```{css, echo=FALSE}
th {
  white-space: nowrap;
}
```

The page presents the mapping between the pilot database fields (see the [DB Schema](db-schema.qmd) page) and the [EFSA Format v2](efsa-format-v2.qmd) columns.

::: {.callout-note}
The location columns this mapping draws on (`dist_to_coast`, the nearest country and
its code, `municipality`, `sea_name`) are **reserved but empty** in the pilot
database. The mapping below is the schema mapping, not a statement about coverage:
the values are derived at the clean stage, which recomputes and overwrites them. See
[DB Schema](./db-schema.qmd).
:::

```{r}
library(dplyr)
library(tibble)

efsa_submission_v2 <- tribble(
  ~Column,              ~Optionality,    ~`Source Table`,   ~`Source Columns`,                             ~`Fixed Text`,  ~Missing, ~Notes,
  "sampleId",           "Recommended",   "sample",          "sample_id",                                    NA,             NA,       NA,
  "samplingStrategy",   "Recommended",   NA,                NA,                                             NA,             "Yes",    "Not recorded in DB",
  "pristine",           "Mandatory",     NA,                NA,                                             "yes",          NA,       NA,
  "collDate",           "Mandatory",     "measurement_time","measurement_date",                             NA,             NA,       NA,
  "placeColl",          "Mandatory",     "survey",          "sea_name",                                     NA,             NA,       NA,
  "lat",                "Recommended",   "survey",          "station_latitude",                             NA,             NA,       NA,
  "lon",                "Recommended",   "survey",          "station_longitude",                            NA,             NA,       NA,
  "sedDepth",           "Mandatory",     "sample",          "layer_upper_boundary + layer_lower_boundary",  NA,             NA,       "Derived: map depth interval to 0-5cm / 5-40cm / >40cm categories",
  "traceEl",            "Mandatory",     "parameter",       "parameter_name",                               NA,             NA,       NA,
  "value",              "Mandatory",     "sediment",        "measured_value",                               NA,             NA,       NA,
  "sd",                 "Optional",      NA,                NA,                                             NA,             "Yes",    "Not recorded in DB",
  "unit",               "Mandatory",     "sediment",        "unit",                                         NA,             NA,       NA,
  "resultType",         "Recommended",   "analysis_method", "measurement_basis",                            NA,             NA,       "Decoded via code_lookup, e.g. D = dry weight",
  "moisture",           "Recommended",   NA,                NA,                                             NA,             "Yes",    "Not recorded in DB",
  "convertedValue",     "Mandatory",     "sediment",        "measured_value + unit",                        NA,             NA,       "Converted to mg/kg dry weight (ug/kg values divided by 1000)",
  "extractionClass",    "Mandatory",     "analysis_method", "chemical_treatment",                           NA,             NA,       "Derived: map chemical treatment code to extraction class",
  "methAn",             "Mandatory",     "analysis_method", "measurement_method_code",                      NA,             NA,       NA,
  "validatedMethod",    "Recommended",   NA,                NA,                                             NA,             "Yes",    "Not recorded in DB",
  "LOD",                "Mandatory",     "lod",             "internal_qa_detection_limit",                  NA,             NA,       NA,
  "LODUnit",            "Mandatory",     "sediment",        "unit",                                         NA,             NA,       "Same unit as the measured value",
  "LOQ",                "Recommended",   "lod",             "internal_qa_quantification_limit",             NA,             NA,       NA,
  "LOQUnit",            "Recommended",   "sediment",        "unit",                                         NA,             NA,       "Same unit as the measured value",
  "accLab",             "Mandatory",     "analysis_method", "accreditation",                                NA,             NA,       NA,
  "phSed",              "Optional",      NA,                NA,                                             NA,             "Yes",    "Not recorded in DB",
  "sieve",              "Mandatory",     "sample",          "matrix",                                       NA,             NA,       "Derived: matrix code indicates sieved fraction (e.g. SED63 = <63µm)",
  "bulkAnalysis",       "Mandatory",     NA,                NA,                                             NA,             "Yes",    "Not recorded in DB",
  "ocSed",              "Mandatory",     NA,                NA,                                             NA,             "Yes",    "Not recorded in DB (no TOC parameter in the dataset)",
  "notes",              "Recommended",   NA,                NA,                                             NA,             "Yes",    "Not recorded in DB"
)

efsa_submission_v2 <- efsa_submission_v2 %>%
  mutate(across(where(is.character), ~coalesce(.x, "")))
```

## Data Mapping for Submission

```{r}
efsa_submission_v2 %>%
  filter(Missing != "Yes") %>%
  select(-Missing)
```

## Full Data Mapping

```{r}
efsa_submission_v2
```

## Data Export Method

As the submission tool is an Excel file, the data must be mapped column by column rather than through a standard row-based export. This approach ensures that the structure of the submission file is not disrupted.
